### 2022-01-19 src-latex/parts/30-dataset.tex
Add annotated example

The annotated named entities are as follows:
\begin{itemize}
    \item PER: a person name or a business name, usually referred to as several person names. Firts names, initials, or civility mentions are included. E.g.: "Alibert (Prosper)", "Allamand frères et Hersent", "Heurtemotte (Vve)", etc.
    \item TITLE: an honorary title, either written or represented by special glyphs. E.g.: "élig.", \scalerel*{\includegraphics{./images/LH.png}}{H} which stand for the Légion d'Honneur, \scalerel*{\includegraphics{./images/GMLondon.png}}{H} which stands for the Great Medal at the London exhibition, .etc.
    \item ACT: the activity of the person (profession or social status) or the type of business. E.g.: "horlogerie", "export en tous genres", "Conseiller d'Etat", "propriétaire", etc.
    \item LOC: a street name or a district name. E.g.: "r. de la Jussienne", "Au Marais", etc.
    \item CARD: a street number, as part of an address. E.g. "16", "5 bis", etc.
    \item FT: a geographic entity type, used to give more details on a location. E.g.: "boutique", "atelier", "fab.", etc.
\end{itemize}

Thus, the extracted corpus contains both the bounding boxes of directory entries, their exact text and the named entities they contain. It can therefore be used to train and improve the entry detection, OCR and NER tools.
% book, page, box, text_ocr, ner_xml (+uid within a page)

>>>>

\bertrand{Show an example for each entity ? Or even a full entry with its entities highlighted ?}
The annotated named entities are as follows:
\begin{itemize}
    \item PER: a person name or a business name, usually referred to as several person names. Firts names, initials, or civility mentions are included. E.g.: "Alibert (Prosper)", "Allamand frères et Hersent", "Heurtemotte (Vve)", etc.
    \item TITLE: an honorary title, either text, glyphs or a combination of the two. E.g: "O. \scalerel*{\includegraphics{./images/LH.png}}{H}" for "Officier de la Légion d'Honneur" or "\scalerel*{\includegraphics{./images/GMLondon.png}}{H}" to represent the Great Medal at the London exhibition.
    \item ACT: the profession or social status of a person summarized in the single concept of activity. E.g.: "horlogerie", "export en tous genres", "Conseiller d'Etat", "propriétaire".
    \item LOC: mostly street names ("r. de la Jussienne"), but may also be neighborhoods ("Marais du Temple") or any indirect spatial references ("au Palais Royal").
    \item CARDINAL: a street number, as part of an address ("16", "5 bis"), or a range of numbers (e.g. "23-25" or "5 à 9").
    \item FT: a geographic entity type, used to give more details on a location, e.g. "boutique", "atelier", "fab." or "dépôt".
\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{images/examples_labeled_entries.pdf}
    \caption{Two manually annotated dictionary entries showing the different labels used in the corpus.}
    \label{fig:my_label}
\end{figure}
Thus, the extracted corpus contains both the bounding boxes of directory entries, their exact text and the named entities they contain. It can therefore be used to train and improve the entry detection, OCR and NER tools.
% book, page, box, text_ocr, ner_xml (+uid within a page)

### 2022-01-19 src-latex/parts/40-ner-approaches.tex
Add annotated example

% !TeX root = ../main-paper.tex
\section{NER Approaches}
We select two deep-learning based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.
We also create an additional CamemBERT model pre-trained on a collection of unannotated directories entries extracted with Pero-OCR.
The NER layer of each three models is then fine-tuned on historical directories as detailed in sections~\ref{subsection:experiment-1} and ~\ref{subsection:experiment-2}.

>>>>

% !TeX root = ../main-paper.tex
\section{NER Approaches}
We select two deep-learning based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.
We also create an additional CamemBERT model pre-trained on a collection of unannotated directories entries extracted with Pero-OCR.
The NER layer of each three models is then fine-tuned on historical directories as detailed in sections~\ref{subsection:experiment-1-setup} and ~\ref{subsection:experiment-2-setup}.

### 2022-01-19 src-latex/parts/50-xp-setup.tex
Add annotated example

\subsection{Experiment 1: NER sensibility to the number of training examples}
\label{subsection:experiment-1}
\bertrand{Find a better name ?}
The constitution of annotated example sets to train a NER model on a target domain a is a critical preliminary step.
Often done manually, possibly from bootstrapped annotations, this task is tedious, time consuming and error prone.
The ability of a model to perform well even with few training examples is a pratical criteria to consider.
In this first experiment, we investigate the performance of the three models when fine-tuned for NER on training sets of decreasing sizes.

@@

All tree models are fine-tuned for NER against every training subset.
The performance metrics are measured against the test set after each training.

@@

\subsection{Experiment 2: NER on entries with OCR noise}
\label{subsection:experiment-2}
The second experiment assesses the impact of the OCR on the quality of the NER results. To this end, we use directory entries whose text has been extracted, corrected, and annotated manually (see Section 3). This reference text is aligned with the text produced on the same entries by Tesseract and PERO OCR. The named entities annotated in the reference text are then projected into the OCRed texts. To make sure that all training, development, and test datasets contain the same annotations, only the entries for which the same number of annotations are found in the reference text and in the OCRised texts are kept. Finally, the reference text and the PERO OCR text are divided into three datasets, using the same method as in Experiment 1. An additional test file is generated from the text produced by Tesseract with the same entries as the test files produced from the reference and PERO OCR texts.

>>>>

\subsection{Experiment 1: NER sensibility to the number of training examples}
\label{subsection:experiment-1-setup}
\bertrand{Find a better name ?}
The constitution of annotated example sets to train a NER model on a target domain a is a critical preliminary step.
Often done manually, possibly from bootstrapped annotations, this task is tedious, time consuming and error prone.
The ability of a model to perform well even with few training examples is a pratical criteria to consider.
In this first experiment, we investigate the performance of the three models when fine-tuned for NER on training sets of decreasing sizes.

@@

All tree models are fine-tuned for NER on each training subset.
The performance metrics are measured against the test set after each training.

@@

\subsection{Experiment 2: Impact of OCR nosie on named entity recognition}
\label{subsection:experiment-2-setup}
Noise introduced by OCR is known to impede the named entity recognition process (see section~\ref{subsection:stoa-ner-on-historical-texts}).
In this second experiment, we are interested in two questions:
\begin{itemize}
    \item which model is the most robust to OCR noise ?
    \item what is most appropriate strategy to fine-tune the NER layer if the texts to label are noisy
\end{itemize}

The second experiment assesses the impact of the OCR on the quality of the NER results. To this end, we use directory entries whose text has been extracted, corrected, and annotated manually (see Section 3). This reference text is aligned with the text produced on the same entries by Tesseract and PERO OCR. The named entities annotated in the reference text are then projected into the OCRed texts. To make sure that all training, development, and test datasets contain the same annotations, only the entries for which the same number of annotations are found in the reference text and in the OCRised texts are kept. Finally, the reference text and the PERO OCR text are divided into three datasets, using the same method as in Experiment 1. An additional test file is generated from the text produced by Tesseract with the same entries as the test files produced from the reference and PERO OCR texts.
